Final Report: Measles risk assessment, modelling, and benefit--cost analyses for New Zealand

### 2015-09-25 draftfinalreport_v4.Rnw
fix some minor grammatical errors

Globally, analyses of 2013 data, the most complete and recent year of data available, suggest measles incidence is highest  and vaccination coverage is lowest in less developed nations (\autoref{table:incidence12} and \autoref{table:cover12}, and \autoref{fig:incidence12} and \autoref{fig:cover12}). Travel of 0--19 year old non-New Zealanders (whether for work, pleasure, etc.) is dominated by people from Australia (\autoref{table:imtravel12} and \autoref{fig:immigration12}). A number of other countries then follow, including United Kingdom, China, Japan and the United States. Travel to Australia is also the most common travel destination (\autoref{table:nztravel12} and \autoref{fig:nztravel12}). Together, these mean that the greatest travel location for New Zealanders and country immigrants and travellers come from is Australia (\autoref{table:immigration12} and \autoref{fig:immigration12}). However, many South Pacific nations also appear in the most visited country list by New Zealanders. Though the precise interactions between these different risk factors are unknown, the most simple, a product of measles incidence in 2013 and traveller numbers in 2013, suggest that those countries with greatest international travel, Australia, UK, China and South East Asian countries, pose the greatest risk to New Zealand for measles importation. Though immigration and travel from and travel to some Asian countries is lower, some Asian countries also pose a high risk of measles importation to New Zealand. Note that South Pacific islands do not appear to pose a greater risk, despite large numbers of travellers of New Zealand origin, perhaps because of low measles incidence in these nations. These data are shown in \autoref{table:risk12}. The breakdown of risk of measles importation by New Zealander and non-New Zealander travellers is in \autoref{table:nzrisk12}, \autoref{table:imrisk12}, \autoref{fig:nzrisk12} and \autoref{fig:imrisk12}. The measles incidence and vaccination coverage for those countries identified as posing the greatest risk for measles importation to New Zealand are in \autoref{table:toprisk}.

@@

The last benefit--cost analysis scenario were performed was one in which more than the minimum proportion of na\"{\i}ve were immunised. In this scenario we did not use the \Sexpr{vacc_reqd_percent}\% of currently na\"{\i}ve population necessary to reach adequate national immunity levels (\autoref{table:coh_vac}), but included additional vaccination so that 50$\%$ of the currently na\"{\i}ve population was immunised, \autoref{table:cost50pc}. In this scenario the additional vaccination reduces the B/C ratio, and the additional vaccines for the extra immunisation mean that the cost per immunised person before the programme becomes costly reduces to \$\Sexpr{break_even_vacc_cost50}  per immunised person \autoref{table:cost50pc}.

@@

Our estimates of the costs of the measles outbreaks in New Zealand suggest that measles management costs per case in New Zealand is high. We used the mean values per case in the absence of alternative data and used all available data. The exact costs will vary, however, our benefit--cost analyses suggest that in all cases catch up immunisation schemes will be financially beneficial for New Zealand (see \autoref{table:cost20} and \autoref{table:cost50}). Additional vaccination beyond the \Sexpr{vacc_reqd_percent}\% of currently na\"{\i}ve population we estimate are required to be vaccinated for $R_0 < 1$ decrease the cost effectiveness of an immunisation program. This is because large outbreaks are avoided once $R_0 < 1$. In the scenario we modelled with 50\% of the currently na\"{\i}ve population immunised, the programme become costly at \$\Sexpr{break_even_vacc_cost50} per immunised person \label{table:cost50pc}. There may be other benefits we have not accounted for in our analyses for additional vaccination. These additional benefits may include reducing the possibility of spatial heterogeneity and future benefits beyond the ten year period we modelled.

@@

Our estimates for the benefit--cost ratio of catch up immunisation are similar to those in other countries. Estimates in Korea suggest catch up immunisation schemes have a benefit--cost ratio of just over one \citep{bae13}. If we presume the $R_0$ estimated for measles in New Zealand is representative and following \Sexpr{vacc_reqd_percent}\% immunisation of the na\"{\i}ve population, measles will be unable to persist ($R_v$ < 1), our B/C ratio is greater than 1. At the national level our B/C ratio is > 1 for either \$50 or \$20 effective immunisation costs. However, the B/C ratios do vary substantially by DHB, with those with smaller populations least cost effective for immunisation programmes. However, the estimated B/C ratio for MMR has been estimated to be between 10.2 and 54.2, much greater than one (see \autoref{table:zhou2}) \citep{zhou4}. As stated above, these authors also consider the costs and benefits of mumps and rubella, so making direct comparison difficult without reanalysis. We did not include the B/C ratio of the complete MMR immunisation programme, though presumably if this was included our B/C ratio may  approached the ratios reported in those analyses that do \citep{zhou4}. Our results are relatively insensitive to changes in immunisation cost, requiring high costs (>\$\Sexpr{break_even_vacc_cost} per immunised person at the national level) before immunisation is predicted to not be beneficial. Similarly, longer discounting periods or 5\% discounting did not affect the B/C ratio substantially (not shown). A spreadsheet is provided for the Ministry to work with.

>>>>

Globally, analyses of 2013 data, the most complete and recent year of data available, suggest measles incidence is highest  and vaccination coverage is lowest in less developed nations (\autoref{table:incidence12} and \autoref{table:cover12}, and \autoref{fig:incidence12} and \autoref{fig:cover12}). Travel of 0--19 year old non-New Zealanders (whether for work, pleasure, etc.) is dominated by people from Australia (\autoref{table:imtravel12} and \autoref{fig:immigration12}). A number of other countries then follow, including United Kingdom, China, Japan and the United States. Travel to Australia is also the most common travel destination (\autoref{table:nztravel12} and \autoref{fig:nztravel12}). Together, these mean that the greatest travel location for New Zealanders and country immigrants and travellers come from is Australia (\autoref{table:immigration12} and \autoref{fig:immigration12}). However, many South Pacific nations also appear in the most visited country list by New Zealanders. Though the precise interactions between these different risk factors are unknown, the most simple, a product of measles incidence in 2013 and traveller numbers in 2013, suggest that those countries with greatest international travel, Australia, UK, China and South East Asian countries, pose the greatest risk to New Zealand for measles importation. Though immigration and travel from some Asian countries is lower, some Asian countries also pose a high risk of measles importation to New Zealand. Note that South Pacific islands do not appear to pose a greater risk, despite large numbers of travellers of New Zealand origin, perhaps because of low measles incidence in these nations. These data are shown in \autoref{table:risk12}. The breakdown of risk of measles importation by New Zealander and non-New Zealander travellers is in \autoref{table:nzrisk12}, \autoref{table:imrisk12}, \autoref{fig:nzrisk12} and \autoref{fig:imrisk12}. The measles incidence and vaccination coverage for those countries identified as posing the greatest risk for measles importation to New Zealand are in \autoref{table:toprisk}.

@@

The last benefit--cost analysis scenario performed was one in which more than the minimum proportion of na\"{\i}ve were immunised. In this scenario we did not use the \Sexpr{vacc_reqd_percent}\% of currently na\"{\i}ve population necessary to reach adequate national immunity levels (\autoref{table:coh_vac}), but included additional vaccination so that 50$\%$ of the currently na\"{\i}ve population was immunised, \autoref{table:cost50pc}. In this scenario the additional vaccination reduces the B/C ratio, and the additional vaccines for the extra immunisation mean that the cost per immunised person before the programme becomes costly reduces to \$\Sexpr{break_even_vacc_cost50}  per immunised person \autoref{table:cost50pc}.

@@

Our estimates of the costs of the measles outbreaks in New Zealand suggest that measles management costs per case in New Zealand is high. We used the mean values per case in the absence of alternative data and used all available data. The exact costs will vary, however, our benefit--cost analyses suggest that in all cases catch up immunisation schemes will be financially beneficial for New Zealand (see \autoref{table:cost20} and \autoref{table:cost50}). Additional vaccination beyond the \Sexpr{vacc_reqd_percent}\% of currently na\"{\i}ve population we estimate are required to be vaccinated for $R_0 < 1$ decrease the cost effectiveness of an immunisation program. This is because large outbreaks are avoided once $R_0 < 1$. In the scenario we modelled with 50\% of the currently na\"{\i}ve population immunised, costs exceed benefits at \$\Sexpr{break_even_vacc_cost50} per immunised person \label{table:cost50pc}. There may be other benefits we have not accounted for in our analyses for additional vaccination. These additional benefits may include reducing the possibility of spatial heterogeneity and future benefits beyond the ten year period we modelled.

@@

Our estimates for the benefit--cost ratio of catch up immunisation are similar to those in other countries. Estimates in Korea suggest catch up immunisation schemes have a benefit--cost ratio of just over one \citep{bae13}. If we presume the $R_0$ estimated for measles in New Zealand is representative and following \Sexpr{vacc_reqd_percent}\% immunisation of the na\"{\i}ve population, measles will be unable to persist ($R_v$ < 1), our B/C ratio is greater than 1. At the national level our B/C ratio is > 1 for either \$50 or \$20 effective immunisation costs. However, the B/C ratios do vary substantially by DHB, with those with smaller populations least cost effective for immunisation programmes. However, the estimated B/C ratio for MMR has been estimated to be between 10.2 and 54.2, much greater than one (see \autoref{table:zhou2}) \citep{zhou4}. As stated above, these authors also consider the costs and benefits of mumps and rubella, so making direct comparison difficult without reanalysis. We did not include the B/C ratio of the complete MMR immunisation programme, though presumably if this was included our B/C ratio may approach the ratios reported in those analyses that do \citep{zhou4}. Our results are relatively insensitive to changes in immunisation cost, requiring high costs (>\$\Sexpr{break_even_vacc_cost} per immunised person at the national level) before immunisation is predicted to not be beneficial. Similarly, longer discounting periods or 5\% discounting did not affect the B/C ratio substantially (not shown). A spreadsheet is provided for the Ministry to work with.
